---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
library(tidygraph, quietly = TRUE)
library(ggraph, quietly = TRUE)
```

# orgsurveyr

The goal of `orgsurveyr` is to facilitate the use of [Thomas Lin Pedersen's](https://www.data-imaginist.com) excellent [`tidygraph`](https://github.com/thomasp85/tidyraph) and [`ggraph`](https://github.com/thomasp85/ggraph) packages with organisational information used by many Human Resources departments, in particular employee surveys.  

## Installation

You can install orgsurveyr from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("ukgovdatascience/orgsurveyr", build_vignettes = TRUE)
```

## Example

An organisation is a very basic type of network known as a tree.  The [`tidygraph`](https://github.com/thomasp85/tidyraph) package lets us simulate and represent a tree structure as follows:

```{r example-1}
tg <- tidygraph::create_tree(13,3) 
tg
```

The [`ggraph`](https://github.com/thomasp85/ggraph) package can be used to plot networks with the familiar [`ggplot2`](http://ggplot2.tidyverse.org) syntax.  Below the simulated organisation is plotted as a dendrogram:

```{r example-2}
  ggraph(tg, 'dendrogram') + geom_edge_diagonal() + 
  geom_node_point(size=5) + theme_bw()
```

The `orgsurveyr` package provides detailed vignettes, convenience functions and example data to help HR analysts make use of the [`ggraph`](https://github.com/thomasp85/ggraph) and [`tidygraph`](https://github.com/thomasp85/tidyraph) packages in the analysis of organisation data.

# Further reading/viewing

- [Data Imaginist](https://www.data-imaginist.com) - Thomas Lin Pedersen's blog
- ['Tidying up your network analysis with tidygraph and ggraph'](https://www.rstudio.com/resources/videos/tidying-up-your-network-analysis-with-tidygraph-and-ggraph/): RStudio::conf presentation from Thomas Lin Pedersen
- [useR 2018 Interview with Thomas Lin Pedersen](https://user2018.r-project.org/blog/2018/04/18/interview-with-thomas-lin-pedersen/)

