---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(tidygraph, quietly = TRUE)
library(ggraph, quietly = TRUE)
```

# orgsurveyr

The goal of `orgsurveyr` is to facilitate the use of Tom Lin Pederson's excellent `tidygrap`h and `ggraph` packages with organisational information used by many Human Resources departments, in particular employee surveys.  

## Installation

You can install orgsurveyr from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("ukgovdatascience/orgsurveyr")
```

## Example

An organisation is a very basic type of network known as a tree.  The tidygraph package lets us simulate and represent a tree structure as follows:

```{r example-1}
tg <- tidygraph::create_tree(13,3) 
tg
```

The `ggraph` package can be used to plot networks with the familiar `ggplot2` syntax.  Below the simulated organisation is plotted as a dendrogram:

```{r example-2}
  ggraph(tg, 'dendrogram') + geom_edge_diagonal() + 
  geom_node_point(size=5) + theme_bw()
```

The `orgsurveyr` provides detailed vignettes, convenience functions and example data to help HR analysts make use of the `ggraph` and `tidygraph` packages in the analysis of organisation data.
