---
title: "Simulating Organisations With tidygraph and ggraph"
author: "Phil Chapman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulating Organisations With tidygraph and ggraph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=5.5,
  fig.height=4
)
library(tidygraph, quietly = TRUE)
library(ggraph, quietly = TRUE)
library(tibble, quietly = TRUE)
library(purrr, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(orgsurveyr, quietly = TRUE)
```

# Introduction

This vignette demonstrates the basic principles of how organisational data can be analysed and visualised using the ggraph and tidygraph packages.  Although the orgsurveyr package provides some convenience functions to facilitate this, it is deliberately kept lightweight with the intention that core ggraph and tidygraph functionality is used instead.  Having established the basic principles behind organisational data analysis, these convenience functions and classes are introduced.

# Simulating The Organisation

## Starting simple

### Creating a simple organisation with tidygraph

A simple tree representing an organisation with three teams, each with three sub-teams can be simulated with `tidygraph` as follows:

```{r}
tg <- tidygraph::create_tree(n = 13, children = 3)
tg
```

The network object is two data frames, one for the nodes and another for the edges.  

### Plotting the simple organisation as a dendrogram
The tidygraph object can be plotted as a dendrogram with `ggraph`

```{r}
ggraph(tg, 'dendrogram') + geom_edge_diagonal() + 
  geom_node_point(shape=21, size=5, fill='white') + theme_bw() 
```

### Manipulating the network with tidygraph

The tidygraph object can be accessed with familiar tidyverse syntax, using the `activate` verb to select either nodes or edges:

```{r}
tg %>% activate(edges) %>% as_tibble()
```

A number of useful functions exist for calculating over and manipulating the network, and these can be combined with the conventional tidyverse verbs such as `mutate` and `filter`.  For example to calculate the depth within the network of each node we can use the `bfs_dist` function:

```{r}
tg <- tg %>%
  mutate(depth = tidygraph::bfs_dist(1))
tg
```

Overlaying this new variable onto the network is as would be expected in any ggplot:

```{r}
ggraph(tg, 'dendrogram') + geom_edge_diagonal() + 
  geom_node_point(aes(fill=depth), shape=21, size=5) + theme_bw() 
```

For more information about how `tidygraph` represents networks see [this blog post](https://www.data-imaginist.com/2017/introducing-tidygraph/).

## Larger organisations

### Simulating a large organisation

To create a larger organisation, we can modify the number nodes in the tree and the maximum number of children per node.

```{r}
bigger_org <- tidygraph::create_tree(n = 40, children = 4)
ggraph(bigger_org, 'dendrogram') + geom_edge_diagonal() + 
  geom_node_point(shape=21, size=2, fill='white') + theme_bw() 

```

### Simulating a symmetrical organisation

To create a more symmetrical organisation, use the formula below:

```{r}
n_children <- 4
max_depth <- 3
regular_n <- sum(n_children^c(0:max_depth))
regular_org <- tidygraph::create_tree(n = regular_n, children = n_children)
ggraph(regular_org, 'dendrogram') + geom_edge_diagonal() + 
  geom_node_point(shape=21, size=2, fill='white') + theme_bw() 
```

The `orgsurvey::create_regular_org` function provides a wrapper around this approach:

```{r}
create_regular_org(n_children=5, max_depth=2) %>%
  ggraph('dendrogram') + geom_edge_diagonal() + 
  geom_node_point(shape=21, size=2, fill='white') + theme_bw() 
```

### Plotting larger organisations

For larger organisations, a circular dendrograph is better as it gives more space for the leaf nodes:

```{r}
ggraph(regular_org, 'dendrogram', circular=TRUE) + geom_edge_diagonal() + 
  geom_node_point(shape=21, size=2, fill='white') + theme_bw() 
```

## Irregular organisations

### General approach

To create a more realistic and less regular, symmetrical organisation, there are two basic approaches: to either build the network up, or trim an existing network down.  In the `orgsurveyr` package the latter approach is taken.  The actions are:

- Build a regular organisation
- Randomly select units to delete
- Propagate deletion status downwards through the organisation
- Delete the unwanted units

### Selecting nodes to delete

To randomly select nodes for deletion, a new variable can be created in the nodes data frame that uses the `rbinom` function to assign a deletion status of true with a probability of 0.3.  Note that the `.N()` function is used to access the nodes data frame within a tidygraph object.

```{r}
set.seed(1234)
tg_random_delete <- create_regular_org(4,2) %>%
  mutate(branch_delete = rbinom(n=nrow(.N()), size=1, prob=0.3))

ggraph(tg_random_delete, 'dendrogram', circular=FALSE) + geom_edge_diagonal() + 
  geom_node_point(aes(fill=as.factor(branch_delete)), shape=21, size=4) + 
  theme_bw() + guides(fill=guide_legend(title = 'Branch Delete'))
```

### Propagating deletion status through the organisation

The next step is to assign all descendent nodes of a node marked as for deletion the same status.  This can be done as follows:

```{r}
tg_random_branch_delete <- tg_random_delete %>%
  mutate(to_delete = tidygraph::map_bfs_dbl(1, .f = function(node, path, ...) {
    max(.N()[c(node, path$node),]$branch_delete)
  }))

tg_random_branch_delete %>%
  ggraph('dendrogram') + geom_edge_diagonal() +
  geom_node_point(aes(fill=as.factor(to_delete)), shape=21, size=4)  + 
  theme_bw() + guides(fill=guide_legend(title = 'To Delete'))
```

The `map_bfs_dbl` function allows us to access the nodes between the current node and the root node and identify whether any already have already been marked for deletion.  The effect can be seen in the plot, where all four of the leaf nodes in the right hand branch have now been marked for deletion.

### Trimming nodes marked for deletion

Deleting nodes is completed using familiar dplyr verbs:

```{r}
tg_random_branch_delete <- tg_random_branch_delete %>%
  filter(to_delete != 1)

#plot
tg_random_branch_delete %>%
  ggraph('dendrogram') + geom_edge_diagonal() +
  geom_node_point(aes(fill=as.factor(to_delete)), shape=21, size=4) + 
  theme_bw() + guides(fill=guide_legend(title = 'To Delete'))
```

### More complex trimming

In the example above, any node was given the same chance of being deleted.  In reality we may wish to have a difference chance of a node being deleted depending on its depth in the organisation.  This can be achieved by first calculating the depth of the organisation using the `bfs_dist` function, and then either applying some function to this or joining another data frame with the required probabilities.  

#### Trimming by a function of depth

The first example demonstrates how simple function can be used to translate depth to likelihood of deletion.  Nodes to be deleted are highlighted in red.

```{r}
set.seed(1234)
tg_trim_by_depth_function <- create_regular_org(4,3) %>%
  mutate(depth = bfs_dist(1),
         prob_deletion = 1-(8-depth)/8,
         branch_delete = rbinom(nrow(.N()),1,prob_deletion),
         to_delete = tidygraph::map_bfs_dbl(1, .f = function(node, path, ...) {
            max(.N()[c(node, path$node),]$branch_delete)
  }))

tg_trim_by_depth_function

tg_trim_by_depth_function %>%
  ggraph('dendrogram', circular=TRUE) + geom_edge_diagonal() +
  geom_node_point(aes(fill=prob_deletion, color=as.factor(to_delete)), 
                  shape=21, size=4, stroke=1) + 
  scale_color_manual(values=c('0'='black', '1'='red')) + 
  theme_bw()


```

#### Trimming by exactly defined probabilities

Second, join to a data frame containing the probabilities defined exactly

```{r}
set.seed(1234)
tg_trim_by_depth_exact <- create_regular_org(4,3) %>%
  mutate(depth = bfs_dist(1)) %>%
  inner_join(data_frame(depth=0:4, 
                        prob_deletion=c(0, 0.1, 0.2, 0.3, 0.4)), by='depth') %>%
  mutate(branch_delete = rbinom(nrow(.N()),1,prob_deletion),
         to_delete = tidygraph::map_bfs_dbl(1, .f = function(node, path, ...) {
            max(.N()[c(node, path$node),]$branch_delete)
  }))

tg_trim_by_depth_exact

tg_trim_by_depth_exact %>%
  ggraph('dendrogram', circular=TRUE) + geom_edge_diagonal() +
  geom_node_point(aes(fill=prob_deletion, color=as.factor(to_delete)), 
                  shape=21, size=4, stroke=1) + 
  scale_color_manual(values=c('0'='black', '1'='red')) + 
  theme_bw()
```

#### Final complex organisation

Either of the above networks can be trimmed and plotted as a reasonable simulation of an organisation:

```{r}
tg_complex <- tg_trim_by_depth_function %>%
  filter(to_delete == 0)

tg_complex %>%
  ggraph('dendrogram', circular=TRUE) + geom_edge_diagonal() +
  geom_node_point(shape=21, size=4, fill='white') + 
  theme_bw()
```

# Simulate the people in the organisation

## Introduction

Having simulated the organisation, we now want to simulate the people within the organisation.  It's the people and what they are doing and saying that we are interested in after all!

## Simulate group sizes

Organisational units within the organisation will have different numbers of people associated with them.  Generally, the leaf nodes will have more people, whereas nodes higher up the organisation will have fewer people directly associated with them.  This is because there will likely only be one manager and perhaps a secretary or a senior person with no further line management responsibility at a non-leaf node.  The leaf nodes, by contrast, will have a number of people with no further line management responsibilities associated with them.

### Define a function to simulate group size

Since we don't want all organisational units to have the same number of people associated with them, let's define a function to simulate the size of the group whether it is a leaf node or not:

```{r}
sim_group_size <- function(is_leaf_node) {
  if (is_leaf_node) {
    ceiling(rbeta(1, 10, 40)*20)
  } else {
    sample(1:3, 1)
  }
}
```

If the group is a leaf node, then the group size will be simulated from the beta distribution, whereas if it isn't then it will be a number between 1 and 3.

### Test the function for simulating group size 

Let's simulate 10000 groups and see how they are distributed, using the `purrr::map` function to apply our function 10,000 times:

```{r}
data_frame(status=rbinom(10000, 1, 0.5)) %>%
  mutate(x=purrr::map_dbl(status, sim_group_size)) %>%
  ggplot(aes(x)) + geom_bar() + facet_wrap(~as.factor(status)) + theme_bw()
```

### Simulate group sizes for the network

The `sim_group_size` function can be applied to the nodes data frame in the `tbl_graph` object as if it were a data frame:

```{r}
tg_group_size <- tg_complex %>%
  mutate(unit_id = dplyr::row_number(),
         is_leaf = node_is_leaf(),
         unit_size = purrr::map_dbl(is_leaf, sim_group_size))

tg_group_size
```

To visualise the group sizes, we can plot them as node labels:

```{r}
tg_group_size %>%
  ggraph('dendrogram', circular=TRUE) + geom_edge_diagonal() +
  geom_node_point(shape=21, size=6, fill='white') +
  geom_node_text(aes(label=unit_size)) + 
  theme_bw()
```

## Simulate the people data

### Simulating individual level variables

Now that we have determined how many people should be in each group, we can simulate them.  Essentially we need a new data frame with one row per person, rather than one row per organisational unit.  There are a variety of ways of doing this, the below example uses purrr::map and tidyr::unnest to simulate an individual level variable with mean of 10 and sd of 3:

```{r}
tg_individuals_df <- tg_group_size %>%
  activate(nodes) %>%
  as_tibble() %>%
  dplyr::select(unit_id, unit_size, depth, is_leaf) %>%
  dplyr::mutate(var1 = purrr::map(unit_size, ~rnorm(., mean=10, sd=3))) %>%
  tidyr::unnest()
tg_individuals_df
```

### Calculating group means for the simulated variable

To plot the individual level data on the network plot, we first summarise it as the group level and then join it back into the nodes data frame of the network:

```{r}
#now calculate means and put info back into network plot
tg_var1_df <- tg_individuals_df %>%
  group_by(unit_id) %>%
  summarise(var1 = mean(var1))

#merge in summarised data
tg_var1 <- tg_group_size %>%
  inner_join(tg_var1_df, by='unit_id')
tg_var1

#plot as a network
tg_var1 %>%
  ggraph('dendrogram', circular=TRUE) + geom_edge_diagonal() +
  geom_node_point(aes(fill=var1), shape=21, size=4) 
```

### Calculating cumulative means for the simulated variable

#### Introduction to cumulative means

Often we don't necessarily want to know the mean for just the people directly associated with an organisational unit, but the mean for all people associated with the current unit _and its children_.  

#### Deriving a new group:individual mapping

What is required is a data frame that maps each unit with all of the units below it in the heirarchy.  This can be achieved using the functionality of the tidygraph::map_bfs_back function in a similar way to how we propagated deletion status through the tree:

```{r}
tg_expanded_df <- tg_group_size %>%
  mutate(child_id = tidygraph::map_bfs_back(unit_id, .f = function(node, path, ...) {
    .N()[c(node, path$node),]$unit_id
  })) %>%
  activate(nodes) %>%
  as_tibble() %>%
  dplyr::transmute(parent_id=unit_id, unit_id=child_id) %>%
  tidyr::unnest() %>%
  arrange(parent_id)
tg_expanded_df
```

### Calculating cumulative variables

We now need to join our individuals data frame with the cumulative mapping data frame, calculate our cumulative variables and then join the result back into the tidygraph object.

First calculate the cumulative variables:

```{r}
tg_cumulative_df <- tg_expanded_df %>%
  inner_join(tg_individuals_df, by='unit_id') %>%
  group_by(parent_id) %>%
  summarise(cumulative_group_size = n(),
            cumulative_var1_mean = mean(var1))
tg_cumulative_df
```

### Merge cumulative variables back into the network

Now merge back into the network object:

```{r}
tg_cumulative <- tg_var1 %>%
  select(unit_id, depth, is_leaf, unit_size, var1) %>%
  inner_join(tg_cumulative_df, by=c('unit_id'='parent_id')) 
tg_cumulative
```

And finally plot - coloured by cumulative mean of the simulated variable, and labelled by cumulative unit size:

```{r}
tg_cumulative %>%
  ggraph('dendrogram', circular=TRUE) + geom_edge_diagonal() +
  geom_node_point(aes(fill=cumulative_var1_mean), shape=21, size=6) +
  geom_node_text(aes(label=cumulative_group_size), color='white', size=3) +
  theme_bw()
```

# Conclusion

In this analysis we have simulated basic organisations, made them more complex by trimming, and then simulated individual level data which we have summarised both at immediate group level and cumulative group level. Simulation of more complex variables is possibly, for example introducing group level effects or adding additional structure into the organisation. 
