---
title: "Setting Up The orgsurveyr Package"
author: "Phil Chapman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setting Up The orgsurveyr Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr, quietly = TRUE)
library(ggplot2, quietly = TRUE)
library(orgsurveyr, quietly = TRUE)

```

# Introduction

This vignette details the steps required to set up and configure the orgsurveyr package, in particular how to use the shiny applications embedded within the package.

# Running shiny apps

## From within RStudio

To run a shiny app within RStudio type the name of the function;

```{r eval=FALSE}
test_viz()
```

More options

- `test_viz` just simulates and organisation
- future viz - permits visualisation of either simulated or real data as long as its formatted in a certain way

## Setting up on a Shiny Server

A detailed explanation of setting up and configuring a Shiny Server is beyond the scope of this vignette, but assuming that a Shiny Server is available and the orgsurveyr package is installed, it is a case of constructing a shiny app file as follows:

```{r eval=FALSE}
library(shiny)
library(orgsurveyr)

shinyApp(ui = test_viz_ui(),
         server = function(input, output) {
           test_viz_server(input, output)
         }
)

```

This can be modified as required and copied to the user's shiny server directory.  

For convenience there is a `make_shiny_app` function which creates a template shiny application file like that above.

# Session Info

```{r}
sessionInfo()
```

