---
title: "Setting Up The orgsurveyr Package"
author: "Phil Chapman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setting Up The orgsurveyr Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr, quietly = TRUE)
library(ggplot2, quietly = TRUE)
library(orgsurveyr, quietly = TRUE)

```

# Introduction

This vignette details the steps required to set up and configure the orgsurveyr package, in particular how to use the shiny applications embedded within the package.

# Running shiny apps within RStudio

## General approach

To run the `test_viz` shiny app within RStudio type the name of the function:

```{r eval=FALSE}
test_viz()
```

Other shiny apps:

- `test_viz` - simulates an organisation with different characteristics
- `orgviz` - permits visualisation of either simulated or real data as long as the data formats described in [orgsurveyr-data-formats](../html/orgsurveyr-data-formats.html) as used.

## Further configuration

Whilst the `test_viz` function has no parameters, the `orgviz` function takes two parameters, `tg` and `df`.  These allow the user to pass a tidygraph object and a data frame containing summary metrics for the organisation into the shiny app for plotting.  For example:

```{r eval=FALSE}
orgviz(tg = orgsurveyr::tg_org, df = orgsurveyr::tg_org_summarised_df)
```

# Running apps on a Shiny Server

## General approach

Assuming that a Shiny Server is available and the orgsurveyr package is installed, it is a case of constructing a shiny app file as follows:

```{r eval=FALSE}
library(shiny)
library(orgsurveyr)

shinyApp(ui = test_viz_ui(),
         server = function(input, output) {
           test_viz_server(input, output)
         }
)

```

This can be modified as required and copied to the user's shiny server directory.  

For convenience there is a [`make_shiny_app`](../html/make_shiny_app.html) function which creates a template shiny application file like that above for both the `test_viz` and `orgviz` shiny apps.

## Further configuration

A shiny app file for the `orgviz` package which includes custom data might look like this:

```{r eval=FALSE}
library(shiny)
library(orgsurveyr)

my_tg <- orgsurveyr::tg_org
my_df <- orgsurveyr::tg_org_summarised_df

shinyApp(ui = orgviz_ui(),
         server = function(input, output) {
           orgviz_server(input, output, tg = my_tg, df = my_df)
         }
)
```

## Docker

The orgsurveyr package can be installed and run within a Docker container that runs both RStudio Server and Shiny Server. This is a convenient way of trying out the Shiny apps with minimal setup overhead and is also a useful way to learn how to configure a Shiny Server. See the README on the package's GitHub repository README for more information.

# Session Info

```{r}
sessionInfo()
```

